---
title: "DSCI 2000 - Assignment"
author: "Dung Ha"
date: "10/19/2020"
output: html_document
---

# Problem 1

Pull out the BLS Release Dates

```{r}
#install.packages("rvest")
#install.packages("dplyr")
library(rvest)
library(dplyr)
source("BLS_Release_Dates.R")
```

Yahoo Finance stock price

```{r}
#install.packages("quantmod")
library(quantmod)
getSymbols("^GSPC", src = "yahoo")
SP_500 <- Cl(GSPC)
rm(GSPC)
SP_500 <- as.data.frame(SP_500)
SP_500 <- tibble::rownames_to_column(SP_500, var = "Date")
SP_500$Date <- as.Date(SP_500$Date)

```

Monthly Unemployment Rate

```{r}
#install.packages("readxl")
library(readxl)
uer <- read_xlsx("SeriesReport-20201006201227_5de657.xlsx")
head(uer)
```

Tidy the unemployment-rates data frame (uer)

```{r}
#Delete the row 1 : 10
uer <- uer[-c(1:10), ]

#Create new data frame
df <- data.frame(DATE = character(), 
                 UNEMPLOYMENT_RATE = numeric())

#Tidy data
for (i in 2:nrow(uer)) {
  for (j in 2:ncol(uer)) {
    date <- paste("01", paste(uer[1, j], uer[i, 1]))
    row <- c(date, uer[i, j])
    names(row) <- c("DATE", "UNEMPLOYMENT_RATE")
    df <- rbind (df, row)
  }
}

#Convert to Date and Numeric
df$DATE <- as.Date(df$DATE, format = "%d %b %Y")
df$UNEMPLOYMENT_RATE <- as.numeric(df$UNEMPLOYMENT_RATE)

#Convert the Date column to Date type
bls$Date <- as.Date(bls$Date, format = "%A, %B %d, %Y")
rm(row, uer, date, i, j)
```

# Problem 2

```{r}
#Only keep the employment situation row, delete other rows
bls <- bls[grepl("Employment Situation for", bls$Release),]

#Adjust the date of df forward a month
#install.packages("lubridate")
library(lubridate)
df$DATE <- df$DATE %m+% months(1)

#Extract the month and yeat of df and bls
df$MONTH <- month(df$DATE)
df$YEAR <- year(df$DATE)
bls$MONTH <- month(bls$Date)
bls$YEAR <- year(bls$Date)

#Match bls and df
df2 <- merge(x = bls, y = df, by = c("YEAR", "MONTH"))

#Match df2 and SP_500
df2 <- merge(x = df2, y = SP_500, by = "Date")

#Delete unnecessary columns and variables
df2 <- df2[,colnames(df2) %in% c("Date", "UNEMPLOYMENT_RATE", "GSPC.Close")]
rm(bls, df, SP_500)
```

# Problem 3

Drawing unemployment rate graph

```{r}
#install.packages("ggplot2")
library (ggplot2)
ggplot(df2, aes(x = Date, y = UNEMPLOYMENT_RATE)) +
  geom_line() +
  theme_bw() +
  xlab("Date") +
  ylab("Unemploymet rate")
```

Drawing closing stock prices of the S&P 500 on the corresponding dates

```{r}
ggplot(df2, aes(x = Date, y = GSPC.Close)) +
  geom_line() +
  theme_bw() +
  xlab("Date") +
  ylab("Closing Stock Price")
```


# Problem 4

The Kendall's t is -0.75 which is close to -1. This means that there is a negative correlation between the unemployment rate and the s&P 500 closing prices on these reporting date.

```{r}
#Sort df2 by date
df2 <- df2[order(df2$Date, decreasing = FALSE),]
t <- cor(df2$UNEMPLOYMENT_RATE, df2$GSPC.Close, method = "kendall", use = "complete.obs")
```
